---
title: "Bike Sales Report"
author: "Tyler Wilson"
date: "7/26/2020"
output: 
  html_document:
    fig_width: 10
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidyquant)
library(timetk)
library(sweep)
library(forecast)
library(patchwork)
library(scales)
library(usmap)

bikes <- bike_sales %>% 
  select(-order.line, -customer.id, -latitude, -longitude, -product.id) %>% 
  mutate_if(is.character, as_factor)

# new trial to create and set new theme
bikes_palette <- setNames(c('#228B22', '#333333'), c('Mountain', 'Road'))

custom_palette <- set_names(hue_pal()(length(unique(bikes$category.secondary))), unique(bikes$category.secondary))

theme_bikes <- ggplot() + scale_fill_manual(values = bikes_palette) +
  theme_minimal()



```

# Build Forcast for Quantity Sold of Primary Categories
This analysis started as a way to learn time series techniques, with out worrying about if they should be used. This analysis is largely based on the vignette produced by RStudio. The vignette can be found at https://cran.rstudio.com/web/packages/sweep/vignettes/SW01_Forecasting_Time_Series_Groups.html. In that example it shows time series predictions for secondary categories. To learn the techniques better and practice with them, I initially decided to use the same data but primary categories.

## Look at Primary Categories
```{r}
# create new dataframe for primary categories
monthly_qty_by_primary <- bike_sales %>% 
  mutate(order.month = as_date(as.yearmon(order.date))) %>% 
  group_by(order.month, category.primary) %>% 
  summarise(total.qty = sum(quantity))

monthly_qty_by_primary %>% 
  ggplot(aes(order.month, total.qty, group = category.primary)) +
  geom_line(aes(color = category.primary)) +
  facet_wrap(~ category.primary) +
  labs(title = 'Quantity Sold: Year Plot',
       x = '',
       y = 'Number Sold',
       subtitle = 'By Primary Category') +
  scale_y_continuous() +
  scale_x_date(date_labels = '%Y') +
  theme_tq() +
  theme(legend.position = 'none')
```

## Create Forecast

### Create Temporary Data Frames
```{r}
# Create time series
monthly_qty_by_primary_ts <- monthly_qty_by_primary %>% 
  group_by(category.primary) %>% 
  nest() %>% 
  mutate(data.ts = map(.x = data,
                       .f = tk_ts,
                       select = -order.month,
                       start = 2011,
                       freq = 12))

monthly_qty_by_primary_ts

# create fit
monthly_qty_by_primary_fit <- monthly_qty_by_primary_ts %>% 
  mutate(fit.ets = map(data.ts, ets, model = 'MAM'))

monthly_qty_by_primary_fit
```

```{r}
monthly_qty_by_primary_fit %>% 
  mutate(glance = map(fit.ets, sw_glance)) %>% 
  unnest(glance)
```

```{r}
prime_augment_fit_ets <- monthly_qty_by_primary_fit %>% 
  mutate(augment = map(fit.ets, sw_augment, timetk_idx = TRUE, rename_index = 'date')) %>% 
  unnest(augment)

prime_augment_fit_ets
```

```{r}
# plot augmented fit residuals (goodness of forecast)
prime_augment_fit_ets %>% 
  ggplot(aes(date, .resid, group = category.primary)) +
  geom_hline(yintercept = 0, color = 'black') +
  geom_line(color = 'red') +
  geom_smooth(method = 'loess') + 
  facet_wrap(~ category.primary, scale = 'free_y') +
  labs(title = 'Bike Quantity Sold by Primary Category',
       subtitle = 'ETS Model Residuals',
       x = '',
       y = 'Residual') +
  scale_x_date(date_labels = '%Y')
```

### Forecast
```{r}
monthly_qty_by_primary_fcast <- monthly_qty_by_primary_fit %>% 
  mutate(fcast.ets = map(fit.ets, forecast, h = 12))

monthly_qty_by_primary_fcast_tidy <- monthly_qty_by_primary_fcast %>% 
  mutate(sweep = map(fcast.ets, sw_sweep, fitted = FALSE, timetk_idx = TRUE)) %>%
  unnest(sweep)

monthly_qty_by_primary_fcast_tidy
```

```{r}
monthly_qty_by_primary_fcast_tidy %>% 
  ggplot(aes(index, total.qty, color = key, group = category.primary)) +
  geom_ribbon(aes(ymin = lo.95,
                  ymax = hi.95),
              fill = 'light blue',
              color = NA) +
  geom_ribbon(aes(ymin = lo.80,
                  ymax = hi.80,
                  fill = key),
              fill = 'blue',
              color = NA,
              alpha = .7) +
  geom_line() +
  facet_wrap(~ category.primary, scales = 'free_y') +
  scale_x_date(date_breaks = '1 year', date_labels = '%Y')
```


## Try normalizing to get better prediction
# ```{r}
# norm_primary <- monthly_qty_by_primary %>% 
#   group_by(category.primary) %>% 
#   mutate(scaled.qty = unlist(scale(total.qty, center = FALSE)[,1])) %>% 
#   ungroup() %>% 
#   select(-total.qty)
# 
# norm_primary %>% 
#   ggplot(aes(order.month, scaled.qty, group = category.primary, color = category.primary)) +
#   geom_line() +
#   facet_wrap(~ category.primary) +
#   labs(title = 'Bike Sales by Primary Category',
#        subtitle = 'Normalized',
#        x = '',
#        y = '') +
#   theme(legend.position = 'none')
# ```
# 
# ```{r}
# # Create time series
# norm_primary_ts <- norm_primary %>% 
#   group_by(category.primary) %>% 
#   nest() %>% 
#   mutate(data.ts = map(.x = data,
#                        .f = tk_ts,
#                        select = -order.month,
#                        start = 2011,
#                        freq = 12))
# 
# norm_primary_ts
# 
# # create fit
# norm_primary_fit <- norm_primary_ts %>% 
#   mutate(fit.ets = map(data.ts, ets))
# 
# norm_primary_fit
# ```
# 
# ```{r}
# norm_primary_fit %>% 
#   mutate(glance = map(fit.ets, sw_glance)) %>% 
#   unnest(glance)
# ```
# 
# ```{r}
# norm_primary_fit_ets <- norm_primary_fit %>% 
#   mutate(augment = map(fit.ets, sw_augment, timetk_idx = TRUE, rename_index = 'date')) %>% 
#   unnest(augment)
# 
# norm_primary_fit_ets
# ```
# 
# ```{r}
# # plot augmented fit residuals (goodness of forecast)
# norm_primary_fit_ets %>% 
#   ggplot(aes(date, .resid, group = category.primary)) +
#   geom_hline(yintercept = 0, color = 'black') +
#   geom_line(color = 'red') +
#   geom_smooth(method = 'loess') + 
#   facet_wrap(~ category.primary, scale = 'free_y') +
#   labs(title = 'Bike Quantity Sold by Primary Category',
#        subtitle = 'ETS Model Residuals',
#        x = '',
#        y = 'Residual') +
#   scale_x_date(date_labels = '%Y')
# ```
# 
# ### Forecast
# ```{r}
# norm_primary_fcast <- norm_primary_fit %>% 
#   mutate(fcast.ets = map(fit.ets, forecast, h = 12))
# 
# norm_primary_fcast_tidy <- norm_primary_fcast %>% 
#   mutate(sweep = map(fcast.ets, sw_sweep, fitted = FALSE, timetk_idx = TRUE)) %>%
#   unnest(sweep)
# 
# norm_primary_fcast_tidy
# ```
# 
# ```{r}
# norm_primary_fcast_tidy %>% 
#   ggplot(aes(index, scaled.qty, color = key, group = category.primary)) +
#   geom_ribbon(aes(ymin = lo.95,
#                   ymax = hi.95),
#               fill = 'light blue',
#               color = NA) +
#   geom_ribbon(aes(ymin = lo.80,
#                   ymax = hi.80,
#                   fill = key),
#               fill = 'blue',
#               color = NA,
#               alpha = .7) +
#   geom_line() +
#   facet_wrap(~ category.primary, scales = 'free_y') +
#   scale_x_date(date_breaks = '1 year', date_labels = '%Y')
# ```

# Bike Sales Exploration
```{r}
bikes <- bike_sales %>% 
  select(-order.line, -customer.id, -latitude, -longitude, -product.id) %>% 
  mutate_if(is.character, as_factor)

bikes
```

```{r}
bikes %>% 
  summary()
```

```{r}
p1 <- bikes %>% 
  group_by(model, category.primary, category.secondary) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity)) %>% 
  slice_max(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = category.primary)) +
  coord_flip() +
  labs(title = 'Top 10 Most Sold Bike Models',
       subtitle = 'By Primary Category',
       x = '',
       y = '',
       fill = '') +
  scale_fill_manual(values = c('#228B22', '#333333'))

p2 <- bikes %>% 
  group_by(model, category.primary, category.secondary) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity)) %>% 
  slice_min(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = category.primary)) +
  coord_flip() +
  labs(title = 'Bottom 10 Bike Models',
       subtitle = 'By Sales and Primary Category',
       x = '',
       y = '',
       fill = '') +
  scale_fill_manual(values = c('#228B22', '#333333'))

p1 + p2
```

```{r}
custom_palette <- set_names(hue_pal()(length(unique(bikes$category.secondary))), unique(bikes$category.secondary))

p3 <- bikes %>% 
  group_by(model, category.primary, category.secondary) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity),
         category.secondary = fct_reorder(category.secondary, quantity)) %>% 
  slice_max(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = category.secondary)) +
  coord_flip() +
  labs(title = 'Top 10 Most Sold Bike Models',
       subtitle = 'By Secondary Category',
       x = '',
       y = '',
       fill = '') +
  scale_fill_manual(values = custom_palette)

p4 <- bikes %>% 
  group_by(model, category.primary, category.secondary) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity),
         category.secondary = fct_reorder(category.secondary, quantity)) %>% 
  slice_min(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = category.secondary)) +
  coord_flip() +
  labs(title = 'Bottom 10 Bike Models',
       subtitle = 'By Sales and Secondary Category',
       x = '',
       y = '',
       fill = '') +
  scale_fill_manual(values = custom_palette)

p3 + p4
```

```{r}
(p1 + p3) / (p2 + p4)
```

```{r}
p5 <- bikes %>% 
  group_by(model, frame) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity)) %>% 
  slice_max(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = frame)) +
  coord_flip() +
  labs(title = 'Top 10 Most Sold Bike Models',
       subtitle = 'By Primary Category',
       x = '',
       y = '',
       fill = '') 

p6 <- bikes %>% 
  group_by(model, frame) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity)) %>% 
  slice_min(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = frame)) +
  coord_flip() +
  labs(title = 'Bottom 10 Bike Models',
       subtitle = 'By Sales and Primary Category',
       x = '',
       y = '',
       fill = '')

p5 + p6
```

```{r}
p7 <- bikes %>% 
  group_by(model, price) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity)) %>% 
  slice_max(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = price)) +
  coord_flip() +
  labs(title = 'Top 10 Most Sold Bike Models',
       subtitle = 'By Primary Category',
       x = '',
       y = '',
       fill = '') +
  scale_fill_gradient(low = '#42bf04', high = '#bf0404', limits = c(0, 13000))

p8 <- bikes %>% 
  group_by(model, price) %>% 
  summarise(quantity = sum(quantity)) %>% 
  arrange(desc(quantity)) %>% 
  ungroup() %>% 
  mutate(model = fct_reorder(model, quantity)) %>% 
  slice_min(model, n = 10) %>% 
  ggplot(aes(model, quantity)) +
  geom_bar(stat = 'identity', 
           aes(fill = price)) +
  coord_flip() +
  labs(title = 'Bottom 10 Bike Models',
       subtitle = 'By Sales and Primary Category',
       x = '',
       y = '',
       fill = '') +
  scale_fill_gradient(low = '#42bf04', high = '#bf0404', limits = c(0, 13000))

p7 + p8
```

```{r}
bikes$fips <- fips(bikes$bikeshop.state)
```

```{r}
map_bikes <- bikes %>% 
  group_by(bikeshop.state, fips) %>% 
  summarise(total_qty = sum(quantity))
  
plot_usmap(data = map_bikes, values = 'total_qty', color = 'dark blue') +
  theme(legend.position = 'right') +
  labs(title = 'Bike Sales',
       subtitle = 'By State') +
  scale_fill_continuous(name = '')
```

## Forecast for Secondary Category

```{r}
# create new dataframe for primary categories
secondary_monthly_qty <- bike_sales %>% 
  mutate(order.month = as_date(as.yearmon(order.date))) %>% 
  group_by(order.month, category.secondary) %>% 
  summarise(total.qty = sum(quantity))

secondary_monthly_qty %>% 
  ggplot(aes(order.month, total.qty, group = category.secondary)) +
  geom_line(aes(color = category.secondary)) +
  facet_wrap(~ category.secondary) +
  labs(title = 'Quantity Sold: Year Plot',
       x = '',
       y = 'Number Sold',
       subtitle = 'By Primary Category') +
  scale_y_continuous() +
  scale_x_date(date_labels = '%Y') +
  theme_tq() +
  theme(legend.position = 'none')
```

## Create Forecast

### Create Temporary Data Frames
```{r}
# Create time series
secondary_monthly_qty_ts <- secondary_monthly_qty %>% 
  group_by(category.secondary) %>% 
  nest() %>% 
  mutate(data.ts = map(.x = data,
                       .f = tk_ts,
                       select = -order.month,
                       start = 2011,
                       freq = 12))

secondary_monthly_qty_ts

# create fit
secondary_monthly_qty_fit <- secondary_monthly_qty_ts %>% 
  mutate(fit.ets = map(data.ts, ets, model = 'MAM'))

secondary_monthly_qty_fit
```

```{r}
secondary_monthly_qty_fit %>% 
  mutate(glance = map(fit.ets, sw_glance)) %>% 
  unnest(glance)
```

```{r}
secondary_augment_fit_ets <- secondary_monthly_qty_fit %>% 
  mutate(augment = map(fit.ets, sw_augment, timetk_idx = TRUE, rename_index = 'date')) %>% 
  unnest(augment)

secondary_augment_fit_ets
```

```{r}
# plot augmented fit residuals (goodness of forecast)
secondary_augment_fit_ets %>% 
  ggplot(aes(date, .resid, group = category.secondary)) +
  geom_hline(yintercept = 0, color = 'black') +
  geom_line(color = 'red') +
  geom_smooth(method = 'loess') + 
  facet_wrap(~ category.secondary, scale = 'free_y') +
  labs(title = 'Bike Quantity Sold by Primary Category',
       subtitle = 'ETS Model Residuals',
       x = '',
       y = 'Residual') +
  scale_x_date(date_labels = '%Y')
```

### Forecast
```{r}
secondary_monthly_qty_fcast <- secondary_monthly_qty_fit %>% 
  mutate(fcast.ets = map(fit.ets, forecast, h = 12))

secondary_monthly_qty_fcast_tidy <- secondary_monthly_qty_fcast %>% 
  mutate(sweep = map(fcast.ets, sw_sweep, fitted = FALSE, timetk_idx = TRUE)) %>%
  unnest(sweep)

secondary_monthly_qty_fcast_tidy
```

```{r}
secondary_monthly_qty_fcast_tidy %>% 
  ggplot(aes(index, total.qty, color = key, group = category.secondary)) +
  geom_ribbon(aes(ymin = lo.95,
                  ymax = hi.95),
              fill = 'light blue',
              color = NA) +
  geom_ribbon(aes(ymin = lo.80,
                  ymax = hi.80,
                  fill = key),
              fill = 'blue',
              color = NA,
              alpha = .7) +
  geom_line() +
  facet_wrap(~ category.secondary, scales = 'free_y') +
  scale_x_date(date_breaks = '1 year', date_labels = '%Y') +
  labs(title = 'Forecast for Secondary Categories',
       x = '',
       y = '')
```


```{r}
secondary_augment_fit_arima <- secondary_monthly_qty_ts %>% 
  mutate(fit_arima = map(data.ts, Arima, order = c(8,0,0)))

secondary_monthly_qty_fcast_arima <- secondary_augment_fit_arima %>% 
  mutate(fcast_arima = map(fit_arima, forecast, h = 12)) %>% 
  mutate(sweep = map(fcast_arima, sw_sweep, fitted = FALSE, timetk_idx = TRUE)) %>%
  unnest(sweep)

secondary_monthly_qty_fcast_arima %>% 
  ggplot(aes(index, total.qty, color = key, group = category.secondary)) +
  geom_ribbon(aes(ymin = lo.95,
                  ymax = hi.95),
              fill = 'light blue',
              color = NA) +
  geom_ribbon(aes(ymin = lo.80,
                  ymax = hi.80,
                  fill = key),
              fill = 'blue',
              color = NA,
              alpha = .7) +
  geom_line() +
  facet_wrap(~ category.secondary, scales = 'free_y') +
  scale_x_date(date_breaks = '1 year', date_labels = '%Y') +
  labs(title = 'Forecast for Secondary Categories',
       x = '',
       y = '')
```

```{r}
secondary_augment_fit_arima <- secondary_monthly_qty_ts %>% 
  mutate(fit_arima = map(data.ts, auto.arima))

secondary_monthly_qty_fcast_arima <- secondary_augment_fit_arima %>% 
  mutate(fcast_arima = map(fit_arima, forecast, h = 12)) %>% 
  mutate(sweep = map(fcast_arima, sw_sweep, fitted = FALSE, timetk_idx = TRUE)) %>%
  unnest(sweep)

secondary_monthly_qty_fcast_arima %>% 
  ggplot(aes(index, total.qty, color = key, group = category.secondary)) +
  geom_ribbon(aes(ymin = lo.95,
                  ymax = hi.95),
              fill = 'light blue',
              color = NA) +
  geom_ribbon(aes(ymin = lo.80,
                  ymax = hi.80,
                  fill = key),
              fill = 'blue',
              color = NA,
              alpha = .7) +
  geom_line() +
  facet_wrap(~ category.secondary, scales = 'free_y') +
  scale_x_date(date_breaks = '1 year', date_labels = '%Y') +
  labs(title = 'Forecast for Secondary Categories',
       x = '',
       y = '')
```

# Revenue
```{r}
bikes %>% 
  group_by(bikeshop.name) %>% 
  summarise(sales = sum(price.ext)) %>% 
  arrange(desc(sales))

sales <- bikes %>% 
  mutate(order.month = as_date(as.yearmon(order.date))) %>% 
  group_by(order.month, category.secondary) %>% 
  summarise(sales = sum(price.ext))
```

```{r}
sales %>% 
  ggplot(aes(order.month, sales, group = category.secondary, color = category.secondary)) +
  geom_line() +
  facet_wrap(~ category.secondary, scales = 'free_y') +
  scale_y_continuous(labels = unit_format(unit = 'M', scale = 1e-6)) +
  theme(legend.position = 'none')
```

```{r}
top_shops <- bikes %>% 
  group_by(bikeshop.name) %>% 
  summarise(sales = sum(price.ext)) %>% 
  slice_max(order_by = sales, n = 10) %>% 
  ungroup() %>% 
  pull(bikeshop.name)


shop_sales <- bikes %>% 
  # filter(bikeshop.name == top_shops) %>% 
  mutate(order.month = as_date(as.yearmon(order.date))) %>% 
  group_by(order.month, bikeshop.name, category.secondary) %>% 
  summarise(sales = sum(price.ext))


```

```{r}
shop_sales %>% 
  ggplot(aes(order.month, sales, group = category.secondary, color = category.secondary)) +
  geom_line() +
  theme(legend.position = 'right') +
  scale_y_continuous(labels = unit_format(unit = 'M', scale = 1e-6)) +
  facet_wrap(~ bikeshop.name, scales = 'free_y')
```

```{r}

```

